---
title: "DetectingBiasinTrafficSearches"
author: "Maggie Meyer"
date: "2023-12-05"
output: html_document
---

```{r setup, include=FALSE}
if(!("pacman" %in% installed.packages()[,"Package"])) install.packages("pacman")
library(pacman)
p_load("tidyverse","rgl", "webshot2", "plotly","reticulate","gganimate","RColorBrewer", "stringr")

path.to.data <- "/Users/maggiemeyer/Dropbox (University of Michigan)/Stanford Open Policing Data/Data for Bias Paper/"
path.to.figures <- "/Users/maggiemeyer/Dropbox (University of Michigan)/Apps/Overleaf/Bias in Searches, Meyer & Gonzalez (2022)/"


D_data <- data.frame(read.csv(paste0(path.to.data, "Durham.csv"), header = T, sep=","))
s_data <- readRDS(paste0(path.to.data, "combined.data.rds"))
local_data <- readRDS(paste0(path.to.data, "non_state_patrol_combined.rds"))

knitr::knit_hooks$set(webgl = hook_webgl)
options(rgl.useNULL = TRUE)
setupKnitr(autoprint = TRUE)
```


```{r, echo=F}
## This section has all the utility functions necessary to produce the plots in our manuscript. 
#1 - two_by_two
##Obtain the 2x2 contingency table for any data frame.
two_by_two <- function(data.frame){
  
  searches = plyr::count(data.frame$search_conducted)[2,2]
  no.searches = plyr::count(data.frame$search_conducted)[1,2]
  contraband = plyr::count(data.frame$contraband_found)[2,2]
  no.contraband = plyr::count(data.frame$contraband_found)[1,2]
  
  tab <- matrix(c(contraband,
                  no.contraband,
                  searches,
                  "x",
                  "y",
                  no.searches,
                  "total contraband",
                  "total no contraband",
                  (searches + no.searches)),ncol=3, byrow=TRUE)
  colnames(tab) <- c('Contraband','No Contraband','Total')
  rownames(tab) <- c('Search','No Search','Total')
  tab <- as.table(tab)
  
  print(tab)
}

#2 - extract.metrics
##Obtain the values from any dataframe that will be used for further computations
extract.metrics <- function(data.frame, name){
  state <- vector()
  
  t_searches = plyr::count(data.frame$search_conducted)[2,2]
  t_no.searches = plyr::count(data.frame$search_conducted)[1,2]
  t_contraband = plyr::count(data.frame$contraband_found)[2,2]
  t_no.contraband = plyr::count(data.frame$contraband_found)[1,2]
  
  white.data = data.frame%>%
    filter(subject_race == "white")
  black.data = data.frame%>%
    filter(subject_race == "black")
  
  w_searches = plyr::count(white.data$search_conducted)[2,2]
  w_no.searches = plyr::count(white.data$search_conducted)[1,2]
  w_contraband = plyr::count(white.data$contraband_found)[2,2]
  w_no.contraband = plyr::count(white.data$contraband_found)[1,2]
  
  b_searches = plyr::count(black.data$search_conducted)[2,2]
  b_no.searches = plyr::count(black.data$search_conducted)[1,2]
  b_contraband = plyr::count(black.data$contraband_found)[2,2]
  b_no.contraband = plyr::count(black.data$contraband_found)[1,2]
  
  PPV.b = b_contraband/b_searches
  PPV.w = w_contraband/w_searches
  PPV = t_contraband/t_searches
  
  S.b = b_searches/(b_searches + b_no.searches)
  S.w = w_searches/(w_searches + w_no.searches)
  S = t_searches/(t_searches + t_no.searches)
  
  metrics <- data.frame(t_searches, t_no.searches, t_contraband, t_no.contraband, PPV, S, 
                        w_searches, w_no.searches, w_contraband, w_no.contraband, PPV.w, S.w, 
                        b_searches, b_no.searches, b_contraband, b_no.contraband, PPV.b, S.b)
  metrics$state <-   substring(name, 1, 2)
  
  #address NAs 
  ifelse(!is.na(metrics),0,metrics)
  return(metrics)
}

#3 - metrics.contraband
##Obtain the full set of feasible contraband and hit rates.
metrics.contraband <- function(data.frame){
  
  contraband.rate <- seq(0, 1, .01)  
  
  b.searches = sum(data.frame$b_searches)
  b.no.searches = sum(data.frame$b_no.searches)
  b.contraband = sum(data.frame$b_contraband)
  b.no.contraband = sum(data.frame$b_no.contraband)
  
  b.total.stops = (b.searches + b.no.searches)
  b.search.rate = (b.searches/b.total.stops)
  b.max = b.no.searches * (b.contraband/b.searches)
  
 #get fa rates 
  p.hit.black <- vector()
  p.fa.black <- vector()
  xvalue.black <- vector()
  search.rate.black <- vector()
  for(i in 1:length(contraband.rate)){
    xvalue.black[i] <- (contraband.rate[i]*b.total.stops - b.contraband)
    p.hit.black[i] <- b.contraband/(b.contraband +xvalue.black[i])
    p.fa.black[i] <- b.no.contraband/(b.no.contraband + (b.no.searches) - xvalue.black[i])
    }
    
  w.searches = sum(data.frame$w_searches) 
  w.no.searches = sum(data.frame$w_no.searches)
  w.contraband = sum(data.frame$w_contraband)
  w.no.contraband = sum(data.frame$w_no.contraband)
  
  w.total.stops = (w.searches + w.no.searches)
  w.search.rate = (w.searches/w.total.stops)
  w.max = w.no.searches * (w.contraband/w.searches)
  
  #sequence
  p.hit.white <- vector()
  p.fa.white <- vector()
  xvalue.white <- vector()
  contraband.rate.white <- vector()
  for(i in 1:length(contraband.rate)){
    xvalue.white[i] <- (contraband.rate[i]*w.total.stops - w.contraband)
    p.hit.white[i] <- w.contraband/(w.contraband +xvalue.white[i])
    p.fa.white[i] <- w.no.contraband/(w.no.contraband + (w.no.searches) - xvalue.white[i])}
    
  bias.fa <- vector()
  bias.hr <- vector()
  bias <-  vector()
  fa.diff <- vector()
  hr.diff <- vector()
  for(i in 1:length(contraband.rate)){
    bias.fa[i] <-  ifelse(p.fa.white[i] < p.fa.black[i], TRUE, FALSE)
    bias.hr[i] <- ifelse(p.hit.black[i] > p.hit.white[i] , TRUE, FALSE)
  bias[i] <- ifelse(bias.fa[i] == TRUE | bias.hr[i] == TRUE, TRUE, FALSE)
  fa.diff[i] <- p.fa.white[i] - p.fa.black[i]
  hr.diff[i] <-  p.hit.white[i] - p.hit.black[i]
  }
  
  
  temp <- data.frame(contraband.rate, xvalue.white, xvalue.black, p.hit.white, p.hit.black, p.fa.white, p.fa.black, w.max, b.max, fa.diff, hr.diff, bias.hr, bias.fa, bias)
  
    for(i in 1:length(contraband.rate)){
    if(1 > temp$xvalue.white[i] |temp$xvalue.white[i] > w.max){temp$p.fa.white[i] <- NA}
    if(1 > temp$xvalue.black[i] |temp$xvalue.black[i] > b.max){temp$p.fa.black[i] <- NA}
    }

  return(temp)
}
```


# Main Analyses

##  Durham 

This section produces the table "Estimating the 2x2 Table in Durham NC" and its corresponding upper bound on x. 
```{r Durham 2x2, echo = FALSE, warning=F}
two_by_two(D_data)
D_two_by_two <- as.numeric(two_by_two(D_data))
x <- D_two_by_two[8]*(D_two_by_two[1]/(D_two_by_two[4]+ D_two_by_two[1]))
x
```

This section produces the tables "Durham NC 2x2 Table by Race" and their corresponding upper bound on x.
```{r, echo=F}
white.data <- D_data %>% filter(subject_race == "white")
two_by_two(white.data)
black.data <- D_data %>% filter(subject_race == "black")
two_by_two(black.data)

#to get the upper bound on x, we use the extract metrics function which does this automatically
D.metrics <- metrics.contraband(extract.metrics(D_data, "NC"))
D.metrics$w.max
D.metrics$b.max
```


### 3D Surfaces

Next we make 3D surfaces of all feasible hit and false alarm rates in Durham NC. This section reproduces the plot Durham NC, Data (Tab. 4): Racial Difference in Hit Rate (left) and False Alarm Rate (right) for all Possible Combinations of x-values. 
```{r Durham 3D, echo=F}
#this is some pre-processing
white.data <- D_data %>% filter(subject_race == "white")
white.search = plyr::count(white.data$search_conducted)[2,2]
white.nosearch = plyr::count(white.data$search_conducted)[1,2]
white.hit = plyr::count(white.data$contraband_found)[2,2]
white.fa = plyr::count(white.data$contraband_found)[1,2] 
w.max = white.nosearch * white.hit/white.search
x.white <- seq(1,w.max-1, length.out=100)
  
black.data <- D_data %>% filter(subject_race == "black")
black.search = plyr::count(black.data$search_conducted)[2,2]
black.nosearch = plyr::count(black.data$search_conducted)[1,2]
black.hit = plyr::count(black.data$contraband_found)[2,2]
black.fa = plyr::count(black.data$contraband_found)[1,2]
b.max = black.nosearch * black.hit/black.search
x.black <- seq(1,b.max-1, length.out=100)

f.bias.hr <- function(x,y) {
  white.hr <- white.hit/(white.hit+x)           
     black.hr <- black.hit/(black.hit+y)
     
     bias <- white.hr - black.hr}
     
hits <- outer(x.white, x.black, f.bias.hr)

f.bias.fa <- function(x,y) {
  white.fa.rate <- white.fa/(white.fa + white.nosearch-x)          
     black.fa.rate <- black.fa/(black.nosearch + black.fa - y)
     
     bias <- white.fa.rate - black.fa.rate}
     
falsealarms <- outer(x.white, x.black, f.bias.fa)

#this generates the plot with the perspective we used in the manuscript
col <- "rgb(255,255,255)"
eye1 <- list(x = -1.64, y = -2.4, z= .8)
#need transpose the matrix to make everything line up
phit <- plot_ly(scene="scene1", z= ~t(hits), z = .28,x = ~x.white, y = ~x.black) %>% 
  add_surface(showscale=FALSE) %>%
  add_surface(z=~matrix(0,100,100),showscale=FALSE, color=I("gray50")) %>% 
  layout(
    scene1 = list(xaxis = list(title = "x white", 
                                    showspikes = FALSE, 
                                    backgroundcolor=col,
                                    showbackground=TRUE,
                                    gridcolor="black",
                                    pagecolor=col), 
                       yaxis = list(title = "x black",
                                    showspikes = FALSE,
                                    backgroundcolor=col,
                                    gridcolor="black",
                                    pagecolor=col,
                                    showbackground=TRUE, autotick=F, tickmode="array", 
                                    tickvals=c(10000,20000,30000,40000)), 
                       zaxis = list(title = "racial difference", 
                                    showspikes = FALSE,
                                    backgroundcolor=col,
                                    showbackground=TRUE,
                                    gridcolor="black",
                                    pagecolor=col, autotick=F, tickmode="array", tickvals=c(.5,0,-.5),range=c(-1,1)), 
                       camera=list(eye=eye1)),
    title = list(text="Hit Rate",y=0.8))

#fa
eye2 <- list(x = 1.55, y = 2.04, z = .28)
eye2 <- list(x = -1.64, y = -2.4, z= .8)

#need transpose the matrix to make everything line up

pfa <- plot_ly(scene="scene2", z= ~t(falsealarms),x = ~x.white, y = ~x.black) %>% add_surface(showscale=FALSE) %>%
  add_surface(z=~matrix(0,nrow(falsealarms),ncol(falsealarms)),showscale=FALSE, color=I("gray50")) %>%
    layout(scene2 = list(
      xaxis = list(title = "x white", 
                   showspikes=FALSE,
                   backgroundcolor=col,
                   showbackground=TRUE,
                   gridcolor="black",
                   pagecolor=col), 
      yaxis = list(title = "x black",showspikes=FALSE,
                   backgroundcolor=col,
                   gridcolor="black",
                   pagecolor=col,
                   showbackground=TRUE, autotick=F, tickmode="array", 
                                    tickvals=c(10000,20000,30000,40000)), 
      zaxis = list(title = "racial difference", 
                   showspikes=FALSE,
                   backgroundcolor=col,
                   showbackground=TRUE,
                   gridcolor="black",
                   pagecolor=col, autotick=F, tickmode="array", tickvals=c(0,-.02,-.04), range=c(-.06,0)), 
      camera=list(eye=eye1)),
    title = list(text="False Alarm Rate",y=0.8),
      paper_bgcolor = col,
      plot_bgcolor = col)

axx <- list(
  gridcolor='rgb(0, 0, 0)',
  zerolinecolor='rgb(0, 0, 0)',
  showbackground=TRUE,
  backgroundcolor='rgb(255, 255, 255)'
)

#do side by side in latex
# save_image(phit, file="./phit.png", scale=2, width=612, height=792)
# save_image(pfa, file="./pfa.png", scale=2, width=612, height=792)

#for html version
#htmlwidgets::saveWidget(as_widget(phit), "phit.html")
#htmlwidgets::saveWidget(as_widget(pfa), "pfa.html")

pfa
phit
```

### Single Blob Plot for Durham NC

This section reproduces the plot "Durham, NC Data (Tab. 4): Graphical Representation of the Overlapping Condition Test"
```{r, echo=F}
temp <- metrics.contraband(extract.metrics(D_data, "NC"))
temp <- temp %>% mutate(minb=min(p.fa.black, na.rm = T),maxb=max(p.fa.black, na.rm = T),minw=min(p.fa.white, na.rm = T),maxw=max(p.fa.white, na.rm = T)) %>% ungroup()

temp$location <- rep("Durham", 101)

durham.blob <- 
ggplot(temp, aes(x=p.fa.white,y=p.fa.black,fill=location)) +
     geom_rect(aes(xmin=minw,xmax=maxw, ymin=minb,ymax=maxb)) + ylim(0,.1) + xlim(0,.1) + geom_abline() + 
  xlab("Feasible FA Rate for White Drivers") + ylab("Feasible FA Rate for Black Drivers")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.background = element_rect(fill='transparent', color=NA),
axis.text = element_text(size = 10))

# ggsave(
#     paste0(path.to.figures,"durham_blob.pdf"),
#     durham.blob,
#     dpi = 1200
#   )

durham.blob
```

### Stop Reasons

This section reproduces the Figure: "Durham, NC Data: Overlapping Condition Test for Stop Reason"
```{r, echo=F}
#create dataframes for each reason
checkpoint <- D_data %>% filter(reason_for_stop == "Checkpoint")
impaired <- D_data %>% filter(reason_for_stop == "Driving While Impaired")
investigation <- D_data %>% filter(reason_for_stop == "Investigation")
other <- D_data %>% filter(reason_for_stop == "Other Motor Vehicle Violation")
safe <- D_data %>% filter(reason_for_stop == "Safe Movement Violation")
seat.belt <- D_data %>% filter(reason_for_stop == "Seat Belt Violation")
speed <- D_data %>% filter(reason_for_stop == "Speed Limit Violation")
stop <- D_data %>% filter(reason_for_stop == "Stop Light/Sign Violation")
equipment <- D_data %>% filter(reason_for_stop == "Vehicle Equipment Violation")
regulatory <- D_data %>% filter(reason_for_stop == "Vehicle Regulatory Violation")


checkpoint.useful <- metrics.contraband(extract.metrics(checkpoint, "checkpoint"))
impaired.useful <- metrics.contraband(extract.metrics(impaired, "impaired"))
investigation.useful <- metrics.contraband(extract.metrics(investigation, "investigation"))
other.useful <- metrics.contraband(extract.metrics(other, "other"))
safe.useful <- metrics.contraband(extract.metrics(safe, "safe"))
seat.belt.useful <- metrics.contraband(extract.metrics(seat.belt, "seat belt"))
speed.useful <- metrics.contraband(extract.metrics(speed, "speed"))
stop.useful <- metrics.contraband(extract.metrics(stop, "stop"))
equipment.useful <- metrics.contraband(extract.metrics(equipment, "equipment"))
regulatory.useful <- metrics.contraband(extract.metrics(regulatory, "regulatory"))

temp <- bind_rows(checkpoint.useful,
                         impaired.useful,
                         investigation.useful,
                         other.useful,
                         safe.useful,
                         seat.belt.useful,
                         speed.useful,
                         stop.useful,
                         equipment.useful,
                         regulatory.useful)

temp$reason <- rep(c("Checkpoint",
                    "Driving While Impaired",
                    "Investigation",
                    "Other Motor Vehicle Violation",
                    "Safe Movement Violation",
                    "Seat Belt Violation",
                    "Speed Limit Violation",
                    "Stop Light/Sign Violation",
                    "Vehicle Equipment Violation",
                    "Vehicle Regulatory Violation"),each = 101)

temp <- temp %>% group_by(reason) %>% mutate(minb=min(p.fa.black, na.rm = T),maxb=max(p.fa.black, na.rm = T),minw=min(p.fa.white, na.rm = T),maxw=max(p.fa.white, na.rm = T)) %>% ungroup()

stop.plot <- ggplot(temp, 
                    aes(x=p.fa.white,y=p.fa.black,fill=reason)) +
  geom_rect(aes(xmin=minw,xmax=maxw, ymin=minb,ymax=maxb)) +
  ylim(0,.2) + 
  xlim(0,.2) + 
  geom_abline() + 
  xlab("Feasible FA Rate for White Drivers") + 
  ylab("Feasible FA Rate for Black Drivers")+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        plot.background = element_rect(fill='transparent', color=NA),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 5),
        legend.key.size = unit(0.5, "cm"),
        legend.title = element_text(size = 10))

# ggsave(
#     paste0(path.to.figures,"stop.pdf"),
#     stop.plot,
#     dpi = 1200
#   )

stop.plot
```


### Search Reasons

This section reproduces the Figure: "Durham, NC Data: Overlapping Condition Test for Search Reason"
```{r, echo=F, warning=F}
#this function has a fixed no search value 
#it should only be used on Durham as it is borrowing the maximal no-search value for Black and White drivers (which is fixed)
metrics.search.reason <- function(data.frame){
  
  contraband.rate <- seq(0, 1, .01)  

  black.data = data.frame%>%
    filter(subject_race == "black")
  
  b.searches = plyr::count(black.data$search_conducted)[1,2]
  b.no.searches = 165469
  b.contraband = plyr::count(black.data$contraband_found)[2,2]
  b.no.contraband = plyr::count(black.data$contraband_found)[1,2]
  
  b.total.stops = (b.searches + b.no.searches)
  b.search.rate = (b.searches/b.total.stops)
  b.max = b.no.searches * (b.contraband/b.searches)

 #get fa rates 
  p.hit.black <- vector()
  p.fa.black <- vector()
  xvalue.black <- vector()
  search.rate.black <- vector()
  for(i in 1:length(contraband.rate)){
    xvalue.black[i] <- (contraband.rate[i]*b.total.stops - b.contraband)
    p.hit.black[i] <- b.contraband/(b.contraband +xvalue.black[i])
    p.fa.black[i] <- b.no.contraband/(b.no.contraband + (b.no.searches) - xvalue.black[i])
    }
  
  white.data = data.frame%>%
  filter(subject_race == "white")
  
  w.searches = plyr::count(white.data$search_conducted)[1,2]
  w.no.searches = 93893
  w.contraband = plyr::count(white.data$contraband_found)[2,2]
  w.no.contraband = plyr::count(white.data$contraband_found)[1,2]
  
  w.total.stops = (w.searches + w.no.searches)
  w.search.rate = (w.searches/w.total.stops)
  w.max = w.no.searches * (w.contraband/w.searches)
  
  #sequence
  p.hit.white <- vector()
  p.fa.white <- vector()
  xvalue.white <- vector()
  contraband.rate.white <- vector()
  for(i in 1:length(contraband.rate)){
    xvalue.white[i] <- (contraband.rate[i]*w.total.stops - w.contraband)
    p.hit.white[i] <- w.contraband/(w.contraband +xvalue.white[i])
    p.fa.white[i] <- w.no.contraband/(w.no.contraband + (w.no.searches) - xvalue.white[i])}
    
  bias.fa <- vector()
  bias.hr <- vector()
  bias <-  vector()
  fa.diff <- vector()
  hr.diff <- vector()
  for(i in 1:length(contraband.rate)){
    bias.fa[i] <-  ifelse(p.fa.white[i] < p.fa.black[i], TRUE, FALSE)
    bias.hr[i] <- ifelse(p.hit.black[i] > p.hit.white[i] , TRUE, FALSE)
  bias[i] <- ifelse(bias.fa[i] == TRUE | bias.hr[i] == TRUE, TRUE, FALSE)
  fa.diff[i] <- p.fa.white[i] - p.fa.black[i]
  hr.diff[i] <-  p.hit.white[i] - p.hit.black[i]
  }
  

  temp <- data.frame(contraband.rate, xvalue.white, xvalue.black, p.hit.white, p.hit.black, p.fa.white, p.fa.black, fa.diff, hr.diff, bias.hr, bias.fa, bias)
  

    for(i in 1:length(contraband.rate)){
    if(1 > temp$xvalue.white[i] |temp$xvalue.white[i] > w.max){temp$p.fa.white[i] <- NA}
    if(1 > temp$xvalue.black[i] |temp$xvalue.black[i] > b.max){temp$p.fa.black[i] <- NA}
    }
  
  return(temp)
}

suspicious.behavior <- metrics.search.reason(D_data %>% filter(str_detect(reason_for_search, "Erratic/Suspicious Behavior")))

informant.tip <- metrics.search.reason(D_data %>% filter(str_detect(reason_for_search, "Informant Tip")))

observation <- metrics.search.reason(D_data %>% filter(str_detect(reason_for_search, "Observation of Suspected Contraband")))

other.info <- metrics.search.reason(D_data %>% filter(str_detect(reason_for_search, "Other Official Information")))

sus.movement <- metrics.search.reason( D_data %>% filter(str_detect(reason_for_search, 'Suspicious Movement')))

witness.obs <- metrics.search.reason(D_data %>% filter(str_detect(reason_for_search, "Witness Observation")))

temp <- bind_rows(suspicious.behavior,
                         informant.tip,
                         observation,
                         other.info,
                         sus.movement,
                         witness.obs)

temp$reason <- rep(c("suspicious behavior",
                     "informant tip",
                     "observation",
                     "other information",
                     "suspicious movement",
                     "witness observation"),each=101)

temp <- temp %>% group_by(reason) %>% mutate(minb=min(p.fa.black, na.rm = T),maxb=max(p.fa.black, na.rm = T),minw=min(p.fa.white, na.rm = T),maxw=max(p.fa.white, na.rm = T)) %>% ungroup()

search.plot <- ggplot(temp, aes(x=p.fa.white,y=p.fa.black,fill=reason)) +
     geom_rect(aes(xmin=minw,xmax=maxw, ymin=minb,ymax=maxb)) + ylim(0,.04) + xlim(0,.04) + geom_abline() + 
  xlab("Feasible FA Rate for White Drivers") + ylab("Feasible FA Rate for Black Drivers")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.background = element_rect(fill='transparent', color=NA),
axis.text = element_text(size = 10))

# ggsave(
#     paste0(path.to.figures,"search.pdf"),
#     search.plot,
#     dpi = 1200
#   )

search.plot
```


## State Analyses

These metrics are presented in the main text body under the section "Analysis of 98 Million Traffic Stops Across 14 States". 
```{r, echo=F}
combined.state.metrics <- metrics.contraband(s_data)
max.whitefa.state <- max(combined.state.metrics$p.fa.white, na.rm=T)
min.blackfa.state <- min(combined.state.metrics$p.fa.black, na.rm=T)

max.whitefa.state
min.blackfa.state
```


This section produces the figure: "US State Patrol Data: Overlapping Condition Test for 14 US States"
```{r, echo=F}
s.metrics <- list()
for(i in s_data$state){
 s.metrics[i] <-list(metrics.contraband(s_data[s_data$state == i,]))
}

az <- s.metrics$az
ca <- s.metrics$ca
co <- s.metrics$co
ct <- s.metrics$ct
ma <- s.metrics$ma
md <- s.metrics$md
mo <- s.metrics$mo
nc <- s.metrics$nc
ri <- s.metrics$ri
sc <- s.metrics$sc
tx <- s.metrics$tx
vt <- s.metrics$vt
wa <- s.metrics$wa
wi <- s.metrics$wi


state.data <- bind_rows(az, ca, co, ct, ma, md, mo, nc, ri, sc, tx, vt, wa, wi)
state.data$state <- rep(c("AZ","CA","CO","CT","MA","MD","MO","NC","RI","SC","TX","VT","WA","WI"),each=101)


state.data <- state.data %>% group_by(state) %>% mutate(minb=min(p.fa.black, na.rm = T),maxb=max(p.fa.black, na.rm = T),minw=min(p.fa.white, na.rm = T),maxw=max(p.fa.white, na.rm = T)) %>% ungroup()

state.plot <- ggplot(state.data,aes(x=p.fa.white,y=p.fa.black,fill=state)) +
  geom_rect(aes(xmin=minw,xmax=maxw, ymin=minb,ymax=maxb)) +
  ylim(0,.3) + 
  xlim(0,.3) + 
  geom_abline() + 
  xlab("Feasible FA Rate for White Drivers") + 
  ylab("Feasible FA Rate for Black Drivers")+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        plot.background = element_rect(fill='transparent', color=NA),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 5),
        legend.key.size = unit(0.5, "cm"),
        legend.title = element_text(size = 10))

# ggsave(
#     paste0(path.to.figures,"state.pdf"),
#     state.plot,
#     dpi = 1200
#   )

state.plot
```



## Insights from the 2x2 Table
```{r, echo=F}
#generate example data
white.hit.o <- 4
white.fa.o <- 23 
white.search.o <- 27
white.nosearch.o <- 73    
white.max.o <- white.nosearch.o * white.hit.o/(white.hit.o + white.fa.o)
  
black.hit.o <- 40
black.fa.o <- 5
black.search.o <- 45
black.nosearch.o <- 55
black.max.o <- black.nosearch.o * black.hit.o/(black.search.o)

x.black.o <- seq(1: black.max.o)
x.white.o <- seq(1: white.max.o)

f.bias.hr.o <- function(x,y) {
  white.hr <- white.hit.o/(white.hit.o+x)           
     black.hr <- black.hit.o/(black.hit.o+y)
     
     bias <- white.hr - black.hr}
hits.o <- outer(x.white.o, x.black.o, f.bias.hr.o)

f.bias.fa.o <- function(x,y) {
  white.fa.rate <- white.fa.o/(white.fa.o + white.nosearch.o-x)          
     black.fa.rate <- black.fa.o/(black.nosearch.o + black.fa.o - y)
     
     bias <- white.fa.rate - black.fa.rate}
falsealarms.o <- outer(x.white.o, x.black.o, f.bias.fa.o)

#hit
col <- "rgb(255,255,255)"
eye1 <- list(x = -2.5, y = 1.4 , z = 1)

phit.o <- plot_ly(scene="scene1", z=~t(hits.o)) %>% 
  add_surface(showscale=FALSE,x = ~x.white.o, y = ~x.black.o) %>%
  add_surface(z=~matrix(0,48,10),showscale=FALSE, color=I("gray50")) %>%
  layout(
    scene1 = list(xaxis = list(title = "x white", 
                                    showspikes = FALSE, 
                                    backgroundcolor=col,
                                    showbackground=TRUE,
                                    gridcolor="black",
                                    pagecolor=col), 
                       yaxis = list(title = "x black",
                                    showspikes = FALSE,
                                    backgroundcolor=col,
                                    gridcolor="black",
                                    pagecolor=col,
                                    showbackground=TRUE), 
                       zaxis = list(title = "racial difference", 
                                    showspikes = FALSE,
                                    backgroundcolor=col,
                                    showbackground=TRUE,
                                    gridcolor="black",
                                    pagecolor=col, autotick=F, tickmode="array", tickvals=c(.3,0,-.5), range=c(-.7,.4)), 
                       camera=list(eye=eye1)),
    title = list(text="Hit Rate",y=0.8))

eye2 <- list(x = -2.5, y = 1.4 , z = 1)

pfa.o <- plot_ly(scene="scene1", z=~t(falsealarms.o)) %>% 
  add_surface(showscale=FALSE,x = ~x.white.o, y = ~x.black.o) %>%
  add_surface(z=~matrix(0,48,10),showscale=FALSE, color=I("gray50")) %>%
  layout(
    scene2 = list(xaxis = list(title = "x white", 
                                    showspikes = FALSE, 
                                    backgroundcolor=col,
                                    showbackground=TRUE,
                                    gridcolor="black",
                                    pagecolor=col), 
                       yaxis = list(title = "x black",
                                    showspikes = FALSE,
                                    backgroundcolor=col,
                                    gridcolor="black",
                                    pagecolor=col,
                                    showbackground=TRUE), 
                       zaxis = list(title = "racial difference", 
                                    showspikes = FALSE,
                                    backgroundcolor=col,
                                    showbackground=TRUE,
                                    gridcolor="black",
                                    pagecolor=col, autotick=F, tickmode="array", tickvals=c(.1,0,-.1), range=c(-.25,.25)), 
                       camera=list(eye=eye2)),
    title = list(text="False Alarm Rate",y=0.8))


# save_image(phit.o, file="./phit.o.png", scale=2, width=612, height=792)
# save_image(pfa.o, file="./pfa.o.png", scale=2, width=612, height=792)

#htmlwidgets::saveWidget(as_widget(phit.o), "phit-o.html")
#htmlwidgets::saveWidget(as_widget(pfa.o), "pfa-o.html")

pfa.o
phit.o
```


# Supplemental Analyses

This section produces the table: "Overlapping Condition Test Table for States"
```{r, echo=F}
s_data <- s_data %>% 
  mutate(w.searchrate = w_searches/(w_searches +w_no.searches),
                                    b.searchrate = b_searches/(b_searches + b_no.searches),
                                    w_PPV = w_contraband/w_searches,
                                    b_PPV = b_contraband/b_searches)
#overlapping condition test
state.data %>%
  select(state, maxw, minb)%>%
  distinct() %>% 
  mutate(maxw = round(maxw, 3), minb = round(minb, 3))

#outcome test
s_data %>% select(state, w_PPV, b_PPV) %>% 
  distinct() %>% 
  mutate(w_PPV = round(w_PPV, 3), b_PPV = round(b_PPV, 3))

#benchmark test
s_data %>% select(state, w.searchrate, b.searchrate) %>% 
  distinct() %>% 
  mutate(w.searchrate = round(w.searchrate, 3), b.searchrate = round(b.searchrate, 3))
```


## Local Municipalities 

This section produces the plot "Overlapping Condition Test for Local Municipalities"
```{r non-state blobs, echo=FALSE}
#Plano has no feasible values of x, appears to be a data reporting issue. 
local_data <- local_data %>%
   filter(state != "plano")

l.metrics <- list()
for(i in local_data$state){
 l.metrics[i] <-list(metrics.contraband(local_data[local_data$state == i,]))
}

oakland <- l.metrics$oakland
san_diego <- l.metrics$san_diego
#san_jose <- l.metrics$san_jose
hartford <- l.metrics$hartford
new_orleans <- l.metrics$new_orleans
charoltte <- l.metrics$charoltte
greensboro <- l.metrics$greensboro
raleigh <- l.metrics$raleigh
winston <- l.metrics$winston
philadelphia <- l.metrics$philadelphia
austin <- l.metrics$austin

#we drop plano because they only have 80 total times they found contraband and all the basis for search are consent searches. 

#we drop san_jose because they have no feasible x-values. Likely issues with data quality and reporting. 

local.data <- bind_rows(oakland, san_diego, hartford, new_orleans, charoltte, greensboro, raleigh, winston, philadelphia, austin)

local.data$location <- rep(c("Oakland", "San Diego", "Hartford", "New Orleans", "Charoltte", "Greensboro", "Raleigh", "Winston", "Philadelphia", "Austin"),each=101)


local.data <- local.data %>% group_by(location) %>% 
  mutate(minb=min(p.fa.black, na.rm = T),
         maxb=max(p.fa.black, na.rm = T),
         minw=min(p.fa.white, na.rm = T),
         maxw=max(p.fa.white, na.rm = T)) %>% ungroup()

local.plot <- 
  ggplot(local.data,aes(x=p.fa.white,y=p.fa.black,fill=location)) +
  geom_rect(aes(xmin=minw,xmax=maxw, ymin=minb,ymax=maxb)) +
  ylim(0,.2) + 
  xlim(0,.2) + 
  geom_abline() + 
  xlab("Feasible FA Rate for White Drivers") + 
  ylab("Feasible FA Rate for Black Drivers")+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        plot.background = element_rect(fill='transparent', color=NA),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 5),
        legend.key.size = unit(0.5, "cm"),
        legend.title = element_text(size = 10))

# ggsave(
#     paste0(path.to.figures,"local_blob.pdf"),
#     local.plot,
#     dpi = 1200
#   )

local.plot
```

This section produces the table "Overlapping Condition Test Table for Local Municipalities"
```{r, echo=F}
local_data <- local_data %>% 
  mutate(w.searchrate = w_searches/(w_searches +w_no.searches),
                                    b.searchrate = b_searches/(b_searches + b_no.searches),
                                    w_PPV = w_contraband/w_searches,
                                    b_PPV = b_contraband/b_searches)

#overlapping condition
local.data %>% dplyr::select(location, maxw, minb)%>% distinct() %>% mutate(maxw = round(maxw, 3),                                                    minb = round(minb, 3))

#outcome
local_data %>% dplyr::select(state, w_PPV, b_PPV)%>% distinct() %>% mutate(w_PPV = round(w_PPV, 3),                                                  b_PPV = round(b_PPV, 3))

#benchmark
local_data %>% dplyr::select(state, w.searchrate, b.searchrate)%>% distinct() %>% mutate(w.searchrate = round(w.searchrate, 3),                                   b.searchrate = round(b.searchrate, 3))

```

